---
title: "AUKUS and the Gravity Framewrok"
author: "Michele Andreoni, Marta Cacopardo, Javier Ospital"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
rm(list =ls())
```

```{r Load Libraries, warning=FALSE, message=FALSE}
library(tidyverse)
library(dplyr)
library(fixest)  
library(readr)  
library(countrycode)
library(ggplot2)
library(scales) 
library(modelsummary)
library(readxl)
```

## 1. Load and Clean GDP Data

```{r Load GDP Data, message=FALSE, warning=FALSE}
gdp_data <- read_csv("/Users/javieroctavioospitalgreslebin/Downloads/WDI_GDP/a4451800-a44a-443a-9309-9013872205ad_Data.csv")

# Clean GDP Data
gdp_data <- gdp_data %>%
  filter(`Series Code` == "NY.GDP.MKTP.CD") %>%
  select(`Country Name`, `Country Code`, matches("\\d{4}")) %>%
  pivot_longer(cols = -c(`Country Name`, `Country Code`), 
               names_to = "Year", 
               values_to = "GDP") %>%
  mutate(Year = as.numeric(str_extract(Year, "\\d{4}")),
         GDP = as.numeric(trimws(GDP))) %>%
  filter(Year >= 2013)
```

## 2. Load and Clean Trade Data

```{r Load Trade Data, message=FALSE, warning=FALSE}
trade_data <- read_csv("/Users/javieroctavioospitalgreslebin/Downloads/Trade_Migration/TradeData_2_26_2025_19_55_48.csv")


trade_data <- trade_data %>%
  select(reporterISO, partnerISO, refYear, flowDesc, fobvalue) %>%  
  filter(reporterISO == "AUS") %>%  
  rename(Year = refYear, Exporter = reporterISO, Importer = partnerISO, Trade_Value = fobvalue) %>%
  mutate(Trade_Value = as.numeric(Trade_Value)) %>%
  filter(!is.na(Trade_Value))
```

## 3. Load and Clean CEPII Data

```{r Load CEPII Data, message=FALSE, warning=FALSE}
gravity_data <- readRDS("/Users/javieroctavioospitalgreslebin/Downloads/Trade_Migration/Gravity_rds_V202211/Gravity_V202211.rds") %>% as_tibble()

gravity_data <- gravity_data %>%
  filter(!is.na(iso3_o) & !is.na(iso3_d)) %>%  
  filter(year == 2021) %>%  
  select(iso3_o, iso3_d, year, dist, distcap, distw_harmonic) %>%
  rename(Exporter = iso3_o, Importer = iso3_d, Distance = dist,
         Capital_Distance = distcap, Pop_Weighted_Distance = distw_harmonic)
```

## 4. Merge Trade, GDP, and CEPII Data

```{r Merge Data, message=FALSE, warning=FALSE}
# select relevant variables 
gdp_data <- gdp_data%>%
  filter(`Country Code` != "WLD")

gravity_data  <- gravity_data %>% drop_na

trade_data <- trade_data %>%
  filter(!Importer %in% c("W00", "X1", "_X"))

trade_gdp <- trade_data %>%
  left_join(gdp_data, by = c("Exporter" = "Country Code", "Year")) %>%
  rename(GDP_Exporter = GDP) %>%
  left_join(gdp_data, by = c("Importer" = "Country Code", "Year")) %>%
  rename(GDP_Importer = GDP) %>%
  left_join(gravity_data, by = c("Exporter", "Importer"))

trade_gdp <- trade_gdp %>% # filter "World", "X1", "_X"
  select(Exporter, Importer, Year, flowDesc, Trade_Value, GDP_Exporter, GDP_Importer, Distance)
#Create log-transformed variables
trade_gdp <- trade_gdp %>%
  mutate(log_GDP_Exporter = log(GDP_Exporter),
         log_GDP_Importer = log(GDP_Importer),
         log_Distance = log(Distance))
```

## **5. Define AUKUS Treatment Variable, AUKUS and China dummies**

```{r Create AUKUS Treatment Variables}
# Define AUKUS partners
aukus_countries <- c("USA", "GBR")

# Create post-AUKUS dummy (1 if 2022 and later, else 0)
trade_gdp <- trade_gdp %>%
  mutate(
    Post_AUKUS = ifelse(Year >= 2022, 1, 0),
    AUKUS_Partner = ifelse(Importer %in% aukus_countries, 1, 0)
  )
```

```{r China dummy, FE }
  	
# Create China-partner dummy
trade_gdp <- trade_gdp %>%
  mutate(
    AUKUS_Partner = ifelse(Importer %in% c("USA", "GBR"), 1, 0),
    China_Partner = ifelse(Importer == "CHN", 1, 0)
  )

# And, we create a country-pair identifier
trade_gdp <- trade_gdp %>%
  mutate(Exporter_Importer = paste0(Exporter, "_", Importer))
```

## **6. Run Gravity Model (PPML) with Country-Pair Fixed Effects**

```{r ppml_model_aukus_china_fe }

ppml_model_aukus_china_fe <- feglm(
  Trade_Value ~ log_GDP_Exporter + log_GDP_Importer + log_Distance + 
    Post_AUKUS * AUKUS_Partner + Post_AUKUS * China_Partner | 
    Exporter + Importer + Year,  # Fixed Effects
  data = trade_gdp, 
  family = poisson
)

summary(ppml_model_aukus_china_fe)
```

## **7. Include Geopolitical Alignment Measures**

```{r Load Geopolitical Data, message=FALSE, warning=FALSE}
ideal_points <- read_csv("/Users/javieroctavioospitalgreslebin/Downloads/Trade_Migration/dataverse_files/IdealpointestimatesAll_Jun2024.csv")
agreement_scores <- read_csv("/Users/javieroctavioospitalgreslebin/Downloads/Trade_Migration/dataverse_files/AgreementScoresAll_Jun2024.csv")

agreement_clean <- agreement_scores %>%
  select(ccode1, ccode2, year, agree) %>%
  rename(Exporter_Code = ccode1, Importer_Code = ccode2, Year = year, Agreement_Score = agree) %>%
  mutate(Exporter = countrycode(Exporter_Code, origin = "cown", destination = "iso3c"),
         Importer = countrycode(Importer_Code, origin = "cown", destination = "iso3c")) %>%
  select(Year, Exporter, Importer, Agreement_Score)
```

```{r Merge Geopolitical Data}
trade_gdp <- trade_gdp %>%
  left_join(agreement_clean, by = c("Exporter", "Importer", "Year")) %>%
  filter(Year < 2024)  # Remove 2024 since no agreement data is available

```

## **8. Run Gravity Model (PPML) with Geopolitical Alignment**

Run Gravity Model (PPML) with Country-Pair Fixed Effects

```{r PPML Model with Geopolitical Agreement}

ppml_model_geo_china <- feglm(
  Trade_Value ~ log_GDP_Exporter + log_GDP_Importer + log_Distance + 
                Post_AUKUS + Agreement_Score + 
                AUKUS_Partner + China_Partner + 
                Post_AUKUS * AUKUS_Partner + 
                Post_AUKUS * China_Partner + 
                Post_AUKUS * Agreement_Score |  
  Exporter + Importer + Year, 
  data = trade_gdp, 
  family = poisson
)
summary(ppml_model_geo_china)

```

### **9.Create Regression Table***


```{r Create Regression Table, warning=FALSE, message=FALSE}
# Save the LaTeX output to a file
modelsummary(
  list( "PPML with FE and Agreement Score" = ppml_model_geo_china),
  output = "/Users/javieroctavioospitalgreslebin/Downloads/Trade_Migration//table.tex",
  title = "PPML Regression Results: AUKUS and Geopolitical Alignment",
  stars = TRUE,
  statistic = "std.error",  # This ensures only standard errors are shown
  gof_map = c("nobs", "r.squared"),
  coef_map = c("log_GDP_Importer" = "log(GDP Importer)",
               "log_Distance" = "log(Distance)",
               "Post_AUKUS" = "Post AUKUS",
               "Agreement_Score" = "Agreement Score",
               "AUKUS_Partner" = "UK and US",
               "China_Partner" = "China",
               "Post_AUKUS:AUKUS_Partner" = "UK and US (post AUKUS)",
               "Post_AUKUS:China_Partner" = "China (post AUKUS)",
               "Post_AUKUS:Agreement_Score" = "Agreement Score (post AUKUS)"))

```






### **10.Making Plots***

```{r Prepare data for visualization, message=FALSE, warning=FALSE, echo=FALSE}

# Ensure Year is numeric
trade_gdp <- trade_gdp %>%
  mutate(Year = as.numeric(Year))

# Aggregate trade by year and country group
trade_trends <- trade_gdp %>%
  group_by(Year, China_Partner, AUKUS_Partner) %>%
  summarise(Trade_Value = sum(Trade_Value, na.rm = TRUE)) %>%
  ungroup()

# Create a new column for group labels
trade_trends <- trade_trends %>%
  mutate(Group = case_when(
    China_Partner == 1 ~ "China",
    AUKUS_Partner == 1 ~ "AUKUS Partners",
    TRUE ~ "Other Countries"
  ))
```


### *10.1 Figure 1: Trade Trends*
```{r Figure 1 Trade Trends, warning=FALSE, echo=FALSE}

# Filter out "Other Countries" and reformat the dataset
trade_trends <- trade_gdp %>%
  filter(AUKUS_Partner == 1 | China_Partner == 1) %>%  # Exclude "Other Countries"
  group_by(Year, China_Partner, AUKUS_Partner) %>%
  summarise(Trade_Value = sum(Trade_Value, na.rm = TRUE)) %>%
  ungroup() %>%
  mutate(Group = case_when(
    China_Partner == 1 ~ "China",
    AUKUS_Partner == 1 ~ "AUKUS Partners"
  ))

# Create the line plot with formatted trade values
ggplot(trade_trends, aes(x = Year, y = Trade_Value, color = Group)) +
  geom_line(size = 1.2) +
  geom_vline(xintercept = 2021, linetype = "dashed", color = "black") +  # Mark AUKUS announcement
  scale_y_continuous(labels = scales::label_number(scale = 1e-9, suffix = "B", accuracy = 0.1)) +  # Convert to Billions (B)
  scale_color_manual(values = c("China" = "#E15759", "AUKUS Partners" = "#4E79A7")) +  # Red for China, Blue for AUKUS
  labs(title = "Trade Trends with AUKUS Partners & China",
       subtitle = "Dashed line marks the announcement of AUKUS in 2021",
       x = "Year", y = "Total Trade Value (Billion EUR)",
       color = "Country Group") +
  theme_minimal()
ggsave("trade trends.pdf", width = 8, height = 6)
```  
  


### *10.2 Figure 2: Change in Trade*

```{r Figure 2: Trade Change Pre/Post AUKUS, warning=FALSE, echo=FALSE}

# Compute percentage change from 2021 to 2022
trade_change <- trade_gdp %>%
  filter(Year %in% c(2021, 2022)) %>%  # Keep only 2021 and 2022
  group_by(Year, China_Partner, AUKUS_Partner) %>%
  summarise(Total_Trade_Value = sum(Trade_Value, na.rm = TRUE), .groups = "drop") %>%
  pivot_wider(names_from = Year, values_from = Total_Trade_Value, names_prefix = "Year_") %>%
  mutate(Trade_Change = (Year_2022 - Year_2021) / Year_2021 * 100,  # Compute % change
         Group = case_when(
           China_Partner == 1 ~ "China",
           AUKUS_Partner == 1 ~ "AUKUS Partners"
         )) %>%
  filter(!is.na(Group))  # Remove NA groups

# Colors: China = Red (#E15759), AUKUS = Blue (#4E79A7)
ggplot(trade_change, aes(x = Group, y = Trade_Change, fill = Group)) +
  geom_bar(stat = "identity", position = "dodge") +
  geom_hline(yintercept = 0, linetype = "dashed", color = "black") +  # Add baseline reference
  scale_fill_manual(values = c("China" = "#E15759", "AUKUS Partners" = "#4E79A7")) +  # Set colors
  labs(title = "Change in Trade from 2021 to 2022",
       subtitle = "Percentage Change in Trade Value (Comparing 2021 to 2022)",
       x = "Country Group", y = "Percentage Change in Trade") +
  theme_minimal() +
  theme(legend.position = "none")  # Remove redundant legend

```


Now, we load data on interventions on trade from China and AUKUS partners on trade with Australia from Global Trade Alert (GTA).
```{r Load Interventions on Trade Data, warning=FALSE, message=FALSE, echo=FALSE}

# Define file path
file_path <- "/Users/javieroctavioospitalgreslebin/Downloads/Trade_Migration/interventions_GTA.xlsx"

# Load the Excel file
gta_data <- read_excel(file_path)

```

And we separate the data the data into Harmful and liberalizing interventions on date. 
```{r Clean Interventions on Trade Data, warning=FALSE, message=FALSE, echo=FALSE}

# Ensure Date Announced is in Date format
gta_data <- gta_data %>%
  mutate(`Date Announced` = ymd(`Date Announced`))

# Filter for the three countries of interest
selected_countries <- c("United Kingdom", "United States of America", "China")

# Separate datasets for Green and Red interventions
gta_green <- gta_data %>%
  filter(`Implementing Jurisdictions` %in% selected_countries, `GTA Evaluation` == "Green") %>%
  mutate(Year = year(`Date Announced`)) %>%
  group_by(Year, `Implementing Jurisdictions`) %>%
  summarise(Num_Interventions = n(), .groups = "drop")

gta_red <- gta_data %>%
  filter(`Implementing Jurisdictions` %in% selected_countries, `GTA Evaluation` == "Red") %>%
  mutate(Year = year(`Date Announced`)) %>%
  group_by(Year, `Implementing Jurisdictions`) %>%
  summarise(Num_Interventions = n(), .groups = "drop")

# Define the AUKUS announcement year
aukus_announcement_year <- 2021
```

### *10.3 Figure 3: Trends on Liberalizing Interventions on Trade*
```{r Figure 3: Liberalizing Interventions, warning=FALSE, message=FALSE, echo=FALSE}

# Create the Green interventions plot (Liberalizing)
p1 <- ggplot(gta_green, aes(x = Year, y = Num_Interventions, color = `Implementing Jurisdictions`)) +
  geom_line(size = 1.2) +
  geom_point(size = 2) +  
  geom_vline(xintercept = aukus_announcement_year, linetype = "dashed", color = "black") + 
  annotate("text", x = aukus_announcement_year + 0.3, y = max(gta_green$Num_Interventions, na.rm = TRUE),
           label = "AUKUS", angle = 0, vjust = 2, hjust = 0.2, size = 4) +
  labs(
       subtitle = "Liberalizing Interventions on Trade with Australia by the UK, US, and China",
       x = "", y = "Number of Interventions",
       color = "" ) +
  theme_minimal() +
  theme(legend.position = "top")
print(p1)

```


### *10.4 Figure 4: Trends on Harmful Interventions on Trade*
```{r Figure 4: Harmful Interventions on trade, warning=FALSE, message=FALSE, echo=FALSE}

# Create the Red interventions plot (Harmful)
p2 <- ggplot(gta_red, aes(x = Year, y = Num_Interventions, color = `Implementing Jurisdictions`)) +
  geom_line(size = 1.2) +
  geom_point(size = 2) +  
  geom_vline(xintercept = aukus_announcement_year, linetype = "dashed", color = "black") + 
  annotate("text", x = aukus_announcement_year + 0.3, y = max(gta_red$Num_Interventions, na.rm = TRUE),
           label = "AUKUS", angle = 0, vjust = 2, hjust = 0.2, size = 4) +
  labs(
       subtitle = "Harmful Interventions on Trade with Australia by the UK, US, and China",
       x = "", y = "Number of Interventions",
       color = "") +
  theme_minimal() +
  theme(legend.position = "top")

# Print the plots
print(p2)
```

### *10.5 Figure 5: Agreement Scores and Geographic Distance*

```{r Figure 5: Agreement Scores and Distance, echo=FALSE}
top60_destinations <- trade_gdp %>%
  filter(Exporter == "AUS", Importer != "W00",, Importer != "_X", flowDesc =="Export")%>%
  group_by(Year)%>%
  slice_max(order_by = Trade_Value, n = 60) %>%
  select(Exporter, Importer, Year, Trade_Value)

top60_origins <- trade_gdp %>%
  filter(Exporter == "AUS", Importer != "W00", Importer != "_X", flowDesc == "Import") %>%
  group_by(Year) %>%
  slice_max(order_by = Trade_Value, n = 60) %>%
  select(Exporter, Importer, Year, Trade_Value, GDP_Importer) %>%
  rename(GDP = GDP_Importer)

merged_top60 <- inner_join(top60_destinations, top60_origins, 
                           by = c("Exporter", "Importer", "Year"),
                           suffix = c("_dest", "_orig")) %>%
  mutate(Trade_Value = Trade_Value_dest + Trade_Value_orig) %>%
  select(Exporter, Importer, Year, Trade_Value, GDP)

dist <- gravity_data %>%
  filter(Exporter == "AUS", Importer != "AUS")%>%
  select(Exporter, Importer, Distance)

# Ensure dist has unique Exporter–Importer combinations
dist_unique <- dist %>% distinct(Exporter, Importer, .keep_all = TRUE)


plots <- list()

for (yr in 2015:2023) {
  
  # Filter data for the current year
  partners_year <- merged_top60 %>% filter(Year == yr)
  geo_allign_year <- agreement_clean %>% filter(Exporter == "AUS", Year == yr)
  
  # Join the trade data with the agreement data (year-specific)
  df_year <- inner_join(partners_year, geo_allign_year, by = c("Exporter", "Importer")) %>%
    select(Exporter, Importer, Trade_Value, Agreement_Score) %>%
    # Join with distance data (which doesn't depend on year)
    inner_join(dist_unique, by = c("Exporter", "Importer")) %>%
    select(Exporter, Importer, Trade_Value, Agreement_Score, Distance) %>%
    # Calculate trade share for the year (sum over all trade values in that year)
    mutate(Trade_Value = Trade_Value / sum(Trade_Value)) %>%
    rename(Australia = Exporter,
           `Trade Partner` = Importer,
           `Trade Share` = Trade_Value,
           `Agreement Score` = Agreement_Score)
  
  # Create the plot for the current year
  p <- ggplot(df_year, aes(x = Distance, y = `Agreement Score`, size =  `Trade Share`,
                             color = case_when(
                               `Trade Partner` == "CHN" ~ "CHN",
                               `Trade Partner` == "USA" ~ "USA",
                               `Trade Partner` == "GBR" ~ "GBR",
                               TRUE ~ "Other"
                             ))) +
    geom_point(shape=16, alpha = 1) +
    scale_color_manual(
      values = c("CHN" = "#E15759", "USA" = "#4E79A7", "GBR"= "#4DAF4A", "Other" = "black"),
      name = "Trade Partners"
    ) +
    scale_size_continuous(range = c(1, 10)) +
    labs(x = "Geographic Distance",
         y = "Agreement Score",
         size = "Trade Volume") +
    ggtitle(paste("Year:", yr)) +
    theme_minimal()
  
  # Store the plot in the list using the year as the name
  plots[[as.character(yr)]] <- p
  
  ggsave(filename = paste0("Trade_Plot_Distance_", yr, ".pdf"), plot = p, width = 8, height = 6)
}

# To view the plot for a specific year (e.g., 2015), you can simply print it:
print(plots[["2015"]])
print(plots[["2023"]])

```

### *10.6 Figure 6: Agreement Scores and GDP*
```{r Figure 6: Agreement Scores and GDP, echo=FALSE}
plots <- list()

for (yr in 2015:2023) {
  
  # Filter data for the current year
  partners_year <- merged_top60 %>% filter(Year == yr)
  geo_allign_year <- agreement_clean %>% filter(Exporter == "AUS", Year == yr)
  
  # Join the trade data with the agreement data (year-specific)
  df_year <- inner_join(partners_year, geo_allign_year, by = c("Exporter", "Importer")) %>%
    select(Exporter, Importer, Trade_Value, Agreement_Score, GDP) %>%
    # Calculate trade share for the year (sum over all trade values in that year)
    mutate(Trade_Value = Trade_Value / sum(Trade_Value)) %>%
    rename(Australia = Exporter,
           `Trade Partner` = Importer,
           `Trade Share` = Trade_Value,
           `Agreement Score` = Agreement_Score)
  
  # Create the plot for the current year
  p <- ggplot(df_year, aes(x = log(GDP), y = `Agreement Score`, size =  `Trade Share`,
                             color = case_when(
                               `Trade Partner` == "CHN" ~ "CHN",
                               `Trade Partner` == "USA" ~ "USA",
                               `Trade Partner` == "GBR" ~ "GBR",
                               TRUE ~ "Other"
                             ))) +
    geom_point(shape=16, alpha = 1) +
    scale_color_manual(
      values = c("CHN" = "#E15759", "USA" = "#4E79A7", "GBR"= "#4DAF4A", "Other" = "black"),
      name = "Trade Partners"
    ) +
    scale_size_continuous(range = c(1, 10)) +
    labs(x = "log(GDP) of Trading Partner",
         y = "Agreement Score",
         size = "Trade Volume") +
    ggtitle(paste("Year:", yr)) +
    theme_minimal()
  
  # Store the plot in the list using the year as the name
  plots[[as.character(yr)]] <- p
  ggsave(filename = paste0("Trade_Plot_GDPs_", yr, ".pdf"), plot = p, width = 8, height = 6)
}

# To view the plot for a specific year (e.g., 2015), you can simply print it:
print(plots[["2015"]])
print(plots[["2023"]])

```

